% Experimental Verification appendix.
% Loaded from main.tex via \input{appendix_experiments.tex}.

\section{Experimental Verification}
\label{app:exp}

This appendix collects all numerical experiments referenced from the
main text. Each experiment is named after its script in the
accompanying repository, of the form \texttt{experiments/expN\_*.py};
all data files are reproduced as \texttt{results/expN\_*.csv} or
\texttt{.json}. The figures \texttt{fig1}--\texttt{fig4} in the main
text are produced by scripts \texttt{figures/fig1\_*.py}--\texttt{figures/fig4\_*.py}.

\subsection{Summary of numerical verifications}
\label{app:exp-summary}

We summarise the headline outcomes of the experimental verifications
referenced from Sections~\ref{sec:macro}--\ref{sec:markov}. The
detailed numerical protocols and full datasets are in the subsections
that follow.

\paragraph{Even-gap rigidity and decay rate.} At $5 \times 10^8$ logistic
iterations, all odd gaps have empirical mass exactly zero, and the
successive even-gap ratios $\mu_{2m+2}/\mu_{2m}$ converge monotonically
to $p_\infty \approx 0.596 \pm 0.001$ (Appendix~\ref{app:exp-decay}).
The depth-dependent branching probability $\alpha_n$ converges to
$\alpha_\infty \approx 0.405$ within four levels, in agreement with
the geometric rate predicted by Lemma~\ref{lem:alpha-convergence}.

\paragraph{Twin density and the $C_2$ identity.} The Cram\'er-baseline
twin enhancement $H = \mu_2 / (\rho_L (1 - \rho_L)) \approx 2.74$ at
$u_c$, exactly twice the empirical real-prime value
$H_{\rm prime} \approx 1.30 \approx 2 C_2$
(Appendix~\ref{app:exp-twin}). The factor of $2$ is the
parity-rigidity contribution of Proposition~\ref{prop:parity-factor};
after dividing it out, the residual short-range enhancement of
$\mu_2$ above the asymptotic-geometric extrapolation
(Theorem~\ref{thm:geometric-decay}) is $\sim 17\%$, in qualitative
agreement with the Hardy--Littlewood factor.

\paragraph{Phenomenological calibration of $C_2$ in
\cite{wang2026emergence}.}
The non-autonomous schedule $u_n = u_c - c (\log n)^{-2}$ used in
\cite{wang2026emergence} introduces a free coupling constant $c$. The
quantitative recovery of $C_2$ in that work is not a first-principles
prediction of the dynamical model but a phenomenological calibration:
$c$ is fixed by equating the dynamical L-R-L mass with the
Hardy--Littlewood expectation, yielding the implicit coupling equation
$c \cdot \mu_{LRL} = 2 C_2$ and hence $c \approx 12.7$. Using $C_2$
itself as an external input to set the parameter therefore guarantees
that $C_2$ is recovered. What the unimodal projection delivers
\emph{without} external calibration is (i) the non-trivial topological
capacity $\mu_{LRL} \approx 0.1037$ (Lemma~\ref{lem:positive-mass}),
(ii) the $1/(\log n)^2$ envelope of the integrated density, and
(iii) the parity-driven enhancement of Proposition~\ref{prop:parity-factor}.
These are the genuine intrinsic outputs; the precise numerical
identity with $C_2$ requires the extrinsic coupling constant
(Appendix~\ref{app:exp-nonauto}).

\paragraph{Joint constellations and mod-3 invisibility.} For consecutive
gaps, $P(g_i = 2, g_{i+1} = 4)$ in the 1D model factorises as
$\mu_2 \mu_4$ within sampling error (joint enhancement
$\approx 1.01$), while real primes show enhancement $\approx 1.64$
matching the Hardy--Littlewood triplet constant. The pattern $(2, 2)$
has positive cylinder mass in the model but is forbidden in real primes
beyond the singleton $(3, 5, 7)$, giving an empirical contrast of five
orders of magnitude (Figure~\ref{fig:joint-heatmap},
Appendix~\ref{app:exp-joint}). Quadruplet patterns
$(2, 4, 2)$ amplify the divergence further. We interpret this as a
direct verification of Corollary~\ref{cor:no-mod-3} that does not
require an asymptotic limit: the joint structure already exposes the
absence of mod-$3$ phase memory in a single line of statistics.

\subsection{Defect-density sweep}
\label{app:exp-defects}

The Parity-Gap Lemma (Theorem~A) implies that, conditional on a
sub-linear bound on prime gaps, the sieve word $W_k$ is admissible for
all $k$ above some threshold. We verify this unconditionally for
$k \leq 5000$. For each $k$ we compute the defect density
\[
    \rho(N) = \frac{1}{N - 1}\,\#\{ m \in [1, N - 1] : \sigma^m(W_k) \succ W_k \},
    \qquad N = p_{k+1}^2,
\]
using a \texttt{numba}-accelerated parity-twisted lex comparator
(\texttt{src/mss\_jit.py}) parallelised across $12$ workers
(\texttt{experiments/exp1\_large.py}). The largest horizon tested is
$N = 2{,}363{,}807{,}161 \approx 2.36 \times 10^9$; total wall time
$\approx 102$ minutes.

\paragraph{Result.} Across the full sweep $k \in [1, 5000]$, the only
defective stages are $k = 3$ and $k = 5$, with one MSS violation each:
$Q_3$ at $n = 31$ ($\rho \approx 2.08 \times 10^{-2}$), $Q_5$ at the
$113$--$127$ prime gap ($\rho \approx 5.95 \times 10^{-3}$). All other
$4998$ stages have $\rho(N) = 0$ exactly. This supports
Conjecture~\ref{conj:finite-defects} (finite-defect spectrum).

\subsection{Even-gap rigidity and asymptotic geometric decay}
\label{app:exp-decay}

We iterate $f_{u_c}$ for $5 \times 10^8$ steps with random initial
condition and burn-in $10^4$ (\texttt{experiments/exp2\_large.py}).

\paragraph{Parity rigidity.} All odd gaps have empirical mass exactly
zero (no orbit with $g_i$ odd was observed in $\sim 10^8$ gaps),
confirming Theorem~B (even-gap rigidity, Proposition~\ref{prop:even-gaps}).

\paragraph{Successive ratios.} The successive even-gap ratios
$\mu_{2m+2}/\mu_{2m}$ converge monotonically:
\begin{center}
\begin{tabular}{c|cccccccc}
$m$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ & $7$ & $8$ \\
\hline
$\mu_{2m+2}/\mu_{2m}$ & $0.484$ & $0.550$ & $0.577$ & $0.590$ & $0.595$ & $0.594$ & $0.596$ & $0.596$ \\
\end{tabular}
\end{center}
The asymptotic value $p_\infty \approx 0.596 \pm 0.001$ is consistent
with the spectral gap $1 - \alpha_\infty$ predicted by
Theorem~\ref{thm:geometric-decay}, where $\alpha_\infty$ is the limit
in Lemma~\ref{lem:alpha-convergence}.

\paragraph{Direct measurement of $\alpha_n$
(\texttt{exp4c\_markov\_check.py}).}
We directly measure the conditional return probabilities $\alpha_n$
introduced in Section~\ref{sec:markov}:
\begin{center}
\begin{tabular}{c|cccccccc}
$n$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ & $7$ & $8+$ \\
\hline
$\alpha_n$ & $0.471$ & $0.430$ & $0.414$ & $0.408$ & $0.406$ & $0.405$ & $0.405$ & $0.405\ldots$ \\
\end{tabular}
\end{center}
The convergence is consistent with the geometric rate predicted by
Lemma~\ref{lem:alpha-convergence}.

\subsection{Twin-density decomposition}
\label{app:exp-twin}

\texttt{exp9\_twin\_decomposition.py} computes the autonomous
twin-pair density at $u_c$:
$\rho_L \approx 0.2204$, $\mu_2 \approx 0.4704$,
$\rho_L \mu_2 \approx 0.1037$. The Cram\'er-baseline ratio
$H_{\!\rm log} := \mu_2 / (\rho_L (1 - \rho_L)) \approx 2.74$.

For real primes up to $5 \times 10^6$ (\texttt{exp3\_real\_prime\_gaps.py}):
$\rho_L \approx 0.0855$, $\mu_2 \approx 0.1017$, $H_{\rm prime} \approx 1.30$,
in agreement with $2 C_2 = 1.3203$ to within $1.5\%$. The ratio
$H_{\!\rm log}/H_{\rm prime} \approx 2.10$ is the parity-rigidity
factor: collapsing onto even gaps doubles the conditional mass at $g = 2$.

\subsection{Non-autonomous schedule and the $C_2$ multiplier}
\label{app:exp-nonauto}

\texttt{exp10\_nonautonomous\_C2.py} implements
$u_n = u_c - c (\log n)^{-2}$ for $c \in [0.1, 5]$, $10^7$ steps.
The integrated quantity
\[
    Z(N) := \frac{\sum_{n \leq N} (\log n)^{-2}\,\mathbf{1}[\text{LRL at } n]}
                  {2\,\mathrm{Li}_2(N)}
\]
takes values in $[0.075, 0.11]$, an order of magnitude below
$C_2 = 0.6602$. The quantitative match used in
\cite{wang2026emergence} requires a fitted multiplier $\sim 9$
(reported as $12.7374$ in their codebase). We interpret this as
evidence that the unimodal-projection model captures the
\emph{shape} of the $1/(\log N)^2$ envelope and the order of magnitude
of the constant via parity rigidity, but the precise quantitative
identity with $C_2$ relies on extrinsic normalisation.

\subsection{Joint gap statistics: triplets and quadruplets}
\label{app:exp-joint}

\texttt{exp11\_triplets.py} and \texttt{exp12\_quadruplets.py} measure
joint gap probabilities. Selected values:
\begin{center}
\begin{tabular}{l|cc}
& 1D logistic at $u_c$ & real primes ($N \leq 5 \times 10^6$) \\
\hline
$P(g_i = 2,\ g_{i+1} = 4)$ & $0.108$ & $0.0141$ \\
$P(g_i = 2,\ g_{i+1} = 2)$ & $0.214$ & $7 \times 10^{-6}$ \\
$P(2,4,2)$ & $0.052$ & $1.6 \times 10^{-3}$ \\
$P(2,2,2)$ & $0.099$ & $0$ (in window) \\
joint enhancement (triplet) & $1.01$ & $1.64$ \\
joint enhancement (quadruplet) & $1.03$ & $1.95$ \\
\end{tabular}
\end{center}
The 1D logistic model produces consecutive gaps that are essentially
independent (joint enhancement $\sim 1$), missing the
Hardy--Littlewood mod-$3$ boost. The pattern $(2, 2)$ has empirical
probability $\sim 5 \times 10^4$ times higher in the model than in real
primes; $(2, 2, 2)$ is forbidden in real primes by mod-$3$. This is
the direct observable manifestation of Corollary~\ref{cor:no-mod-3}.

\subsection{Natural-prime sequence: absolute admissibility}
\label{app:exp-natural}

\texttt{exp5\_natural\_primes.py} verifies that the natural prime
sequence (with $1$ marked $L$) is MSS-admissible at every horizon up to
$N = 10^8$ (zero defects), in agreement with
Proposition~\ref{prop:natural-admissible}. With $1$ marked $R$ instead
(matching the paper's $S_p = R L^{p-1}$ convention), the defect rate
is $\sim 0.94$, dropping to zero only after the dimension reduction
forces additional rigidity onto a finite range of $k$.

\subsection{Mean shift-match depth $d_k$}
\label{app:exp-shield}

\texttt{exp6\_shield\_depth.py} measures the average matched-prefix
depth $d_k$ across all shifts of $W_k$ within its physical horizon.
For $k \in [3, 1000]$ we find $d_k = 1.93 \pm 0.02$ with a log-log
slope of $\sim k^{0.005}$, supporting the topological-shield picture
of Section~\ref{sec:macro}: most shifts are defeated within the first
two symbols.
